Statistical Inference for inter-arrival times of extreme events in bursty time series

### 2018-10-30 article_springer/article_springer.Rmd
Restructurings, new title, Data Sexample section

---
title: "Peaks Over Threshold for Bursty Time Series"
# titlerunning: "CTRE: POT for Bursty Time Series"
thanks: |
    Peter Straka was supported by the 
    Discovery Early Career Research Award DE160101147 on the Project 
    "Predicting Extremes when Events Occur in Bursts" by the Australian
    Research Council.
    Katharina Hees was supported by 
    the DAAD co-financed by the German Federal Ministry of Education and 
    Research (BMBF).

authors: 
- name: "Katharina Hees"
  address: University of Dortmund
  email: hees@statistik.uni-dortmund.de

@@

bibliography: CTRMstats.bib

output: 
  rticles::springer_article: 
    fig_caption: yes

@@

preamble: |
  \usepackage{amssymb}
---

```{r get-rticles, eval=FALSE, include=FALSE, cache=TRUE}
# use this rticles version: 
devtools::install_github("rstudio/rticles")
# use the newest MittagLeffleR version:
devtools::install_github("UNSW-MATH/MittagLeffleR")
library(MittagLeffleR)
devtools::install_github("UNSW-MATH/CTRE")
library(CTRE)
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = TRUE,
  message = FALSE,
  fig.height = 3, out.width = '\\textwidth'
  )
library(CTRE)
library(MittagLeffleR)
library(magrittr)
```

@@

We review the marked renewal process in Section 2, and 
derive a scaling limit theorem for inter-exceedance times in Section 3.
We give a statistical procedure to estimate model parameters via stability plots 
in Section 5, but to set the stage we first discuss inference for the 
Mittag-Leffler distribution in Section 4. 
(A simulation study of the effectiveness of our statistical procedure is 
given in the appendix.)
Diagnostic plots for model criticism are discussed in Section 6. 
In Section 7, we discuss the memory property of the 
Mittag-Leffler distribution, and how it affects the predictive distribution 
for the time until the next threshold crossing event. 
Section 8 concludes. 
For all statistical computations we have used R [@R]. 
All code and data used for the analysis in this article has been organized into 
an R package `CTRE` (<https://github.com/UNSW-MATH/CTRE>).
The source code for the figures generated in this manuscript is 
available online at <https://github.com/UNSW-MATH/bursty-POT>.

@@

As a model for extreme observations, we use a Marked Renewal Process (MRP):

@@

: Let $(W,J), (W_1, J_1), (W_2, J_2), \ldots$ be i.i.d. pairs of random 
variables, where the $W_k > 0$ are interpreted as the *waiting times* 
and $J_k \in [x_L, x_R]$ as the *event magnitudes* 
($x_L \in [-\infty, +\infty), x_R \in (-\infty, +\infty]$). 
If $W$ and $J$ are independent, the Marked Renewal Process is said to be *uncoupled*. 
\qed

@@

: Given a threshold $\ell \in (x_L, x_R)$, 
consider the stopping time 
$$\tau(\ell) := \min\{k: J_k > \ell\},\quad \ell \in (x_L, x_R).$$
Define the pair of random variables $(X(\ell), T(\ell))$ via 
$$X(\ell) = J_{\tau(\ell)} - \ell, \quad 
T(\ell) = \sum_{k=1}^{\tau(\ell)} W_k.$$
By restarting the MRP at $\tau(\ell)$, inductively define the
two i.i.d. sequences $T(\ell,n)$ and $X(\ell, n)$, $n \in \mathbb N$, 
called the "interarrival times" and the "exceedances", respectively.
The pair sequence $(T(\ell, n), W(\ell, n))_{n \in \mathbb N}$ is called 
a Continuous Time Random Exceedance model (CTRE). 
If the underlying MRP is uncoupled, then the CTRE is also called 
uncoupled. 
\qed

```{r thresholdedBursty, message=FALSE, fig.height=7, fig.width=9, fig.cap="\\label{fig:thresholdedBursty}Exceedances (red) and Times until Exceedance (durations between blue crosses) for a given threshold $\\ell$ (dashed line)."}
tail <- params$tail
n <- params$n 
sigma <- (cos(pi*tail/2))^(1/tail)
times <- cumsum(stabledist::rstable(n, tail, 1, sigma, pm=1))
#times<-cumsum(rpareto(n,tail))
magnitudes <- extRemes::revd(n, scale = 1, shape = 0)
sim_ctre <- ctre(data.frame(times, magnitudes)) 
par(mfrow = c(2,1))
plot(sim_ctre, p = 0.01, main = "Simulated MRP") 
flares %>% ctre() %>% plot(p = 0.02, log = 'y', main = "HXRBS data") 
```

In this article, we restrict ourselves to the uncoupled case, where
the two sequences $X(\ell, n)_{n \in \mathbb N}$
and $T(\ell, n)_{n \in \mathbb N}$ are independent.[^1]
Figure \ref{fig:thresholdedBursty} shows a simulated dataset in the top 
panel, where $W$ has a stable distribution with tail parameter 
$\beta =$ `r params$tail` (and skewness $1$ and location $0$), and 
where $J$ is from a standard Gumbel distribution. In the bottom panel, 
we plot a time series of solar flare intensities derived from a NASA
dataset [@HXRBS][^*]. Clearly, the simulated data exhibit long intervals 
_without any_ events, whereas in the real-world dataset events appear 
continuously. The threshold exceedances, however, appear to have similar
statistical behaviour in both models. Observations below a threshold are 
commonly discarded in Extreme Value Theory (POT approach); likewise, the CTRE model
interprets these observations as noise and filters them out.

[^1]: To see why, note that $X(\ell)$ is, in distribution, simply equal to 
  $J | J > \ell$, independent of any waiting time $W_k$.

[^*]: The "complete Hard X Ray Burst Spectrometer event list" 
is a comprehensive reference for all measurements of the Hard X 
Ray Burst Spectrometer on NASA's Solar Maximum Mission from the time of 
launch on Feb 14, 1980 to the end of the mission in Dec 1989. 12,776 
events were detected, with the "vast 
majority being solar flares". The list includes the start time, peak time, 
duration, and peak rate of each event. We have used "start time" as the 
variable for event times, and "peak rate" as the variable for event 
magnitudes.

@@

: Let the waiting times $J_k$ be in the domain of attraction of a 
positively skewed sum-stable law with stability parameter $0 < \beta < 1$; 
more precisely,
\begin{align} \label{eq:stability}
\frac{W_1 + \ldots + W_n}{b(n)} \overset{d}{\longrightarrow} D, 
\quad n \to \infty
\end{align}
for a function $b(n)$ which is regularly varying at $\infty$ with 
parameter $1/\beta$, and where $\mathbf E[\exp(-sD)] = \exp(-s^\beta)$. 
Write $p := \mathbf P(J > \ell)$. Then the weak convergence
$$
\frac{T(\ell)} {b(1/p)} \to W_\beta \quad \text{ as } \quad \ell \uparrow x_R
$$
holds, where the Mittag-Leffler random variable $W_\beta$ is defined on 
the positive real numbers via 
$$
\mathbf E[\exp(-sW_\beta)] = \frac{1}{1+s^\beta}.
$$
\qed

\noindent For a scale parameter $\sigma > 0$, we write 
${\rm ML}(\beta, \sigma)$ for the distribution of $\sigma W_\beta$. 
The Mittag-Leffler distribution with parameter $\beta \in (0,1]$ is 
a heavy-tailed positive distribution for $\beta < 1$, with infinite mean. 
However, as $\beta \uparrow 1$, ${\rm ML}(\beta, \sigma)$ converges 
weakly to the exponential distribution ${\rm Exp}(\sigma)$.
This means that although its moments are all infinite, the Mittag-Leffler
distribution may (if $\beta$ is close to 1) be indistinguishable from the 
exponential distribution, for the purposes of applied statistics. 
For a detailed reference on the Mittag-Leffler distribution, see e.g. 
@Haubold11, and for algorithms, see e.g. the R package `MittagLeffleR`
[@MittagLeffleR].

#### _Proof of Theorem:_
We interpret the threshold crossing time $T(\ell)$ as the hitting time of the 
underlying CTRM (Continuous Time Random Maxima) or "max-renewal process", 
and then utilize a result by @MeerschaertStoev08. 
The running maximum process is defined as 
$$
M(c) := J_1 \vee \ldots \vee J_{\lfloor c \rfloor},
$$
and since we assume that the $J_k$ have a continuous distribution, there exist 
norming functions $a(c)$ and $d(c)$ such that 
$$
\mathbf P\left[ \frac{M(c) - d(c)}{a(c)} \le \ell^* \right] 
\longrightarrow F(\ell^*), \quad t \to \infty
$$
where $F$ is a generalized extreme value distribution, and $\ell^*$ is any 
value from the support of $F$. 
The CTRM process is then defined via 
$$
V(t) = M(N(t)), \quad t \ge 0
$$
where $N(t)$ is the renewal process associated with the waiting times 
$W_k$: 
$$
N(t) = \max\{n: W_1 + \ldots + W_n \le t\}.
$$
Now a key observation is that
$$
T(\ell) = \inf\{t: V(t) > \ell\}, 
$$
and that 
$$
T(\ell) > t \quad \text{ if and only if } \quad V(t) \le \ell.
$$
By [Theorem 3.1, @MeerschaertStoev08], we have the stochastic process 
convergence 
$$
\frac{V(ct) - d(\tilde b(c))}{a(\tilde b(c))} 
\stackrel{d}{\longrightarrow} Y(t), \quad t > 0.
$$
where $Y(t)$ is a time-changed ("subordinated") extremal process, and where 
$\tilde b(c)$ is a regularly varying norming function which is
_inverse_ to $b(c)$, in the sense that 
$b(\tilde b(c)) \sim c \sim \tilde b(b(c))$.

@@

**Remark:**
: If $\beta = 1$, the result of the Theorem above is standard, see e.g. 
Equation (2.2) in @Gut1999.

@@

```{r fig:QQ-plots, fig.cap="Pareto vs. Mittag-Leffler distribution, both with same tail exponent 0.8. Top row: solar flare magnitudes data, fitting nicely with a Pareto distribution. Bottom row: simulated Mittag-Leffler data. Left column: static QQ-Estimator plots, assessing a fit against the Pareto distribution. Centre column: dynamic QQ-Estimator plots. Right column: a usual QQ-Plot based on Mittag-Leffler quantiles.\\label{fig:QQ-plots}", fig.height=5.5, fig.width=7}
par(mfrow = c(2,3))

flare_mags <- actuar::rpareto(n = 500, shape = 0.8, scale = 1)
flares_exponent <- flare_mags %>% qqestplot_static()

flare_mags %>% tea::qqestplot(conf.int = TRUE)
abline(h = 0.8, lty = 2, col = 2)

flare_mags %>% mlqqplot(tail = 0.8, log = 'xy')
(flare_mags %>% MittagLeffleR::logMomentEstimator())[1:2] -> MLparams
flare_mags %>% qqline(distribution = function(p) qml(p, tail = 0.8, scale = MLparams[2]))

ML_data <- MittagLeffleR::rml(500, 0.8)
ML_exponent <- ML_data %>% qqestplot_static()

ML_exponent2 <- ML_data %>% tea::qqestplot(conf.int = TRUE)
abline(h=0.8, lty = 2, col = 2)

ML_data %>% mlqqplot(0.8, log = 'xy')
qqline(ML_data, distribution = function(p) qml(p, 0.8), col = 2)

```

The first method proposed for parameter estimation of the Mittag-Leffler distribution was the fractional moment estimator by [@kozubowski2001]. One drawback of this method is that constant priors for the tail parameter are needed for the calculation of the estimates. [@Cahoy2010] proposed an estimator based on the log-Moments which doesn't need prior knowledge. Furthermore, they showed in a simulation study that he log-Moments estimator outclasses the fractional moment estimator regarding bias and root mean squared error (RMSE). Due to the form of the density function there exists no closed form for the Maximum Likelihood estimator (MLE). In the RPackage `MittagLeffleR` Maximum Likelihood estimation is implemented via numerical optimization. The MLE outperforms the log-ent estimator regarding bias and RMSE for big enough sample sizes, but is extremely computational intensive. In figure XX one can see that both mehtods work quite well even for very low sample sizes.    
Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators such as the Hill estimator. However, these methods are less statistically efficient since they only use a portion of the information contained in the data, as was already mentionded by [@kozubowski2001]. Addtionally they need a specification of a tuning parameter.

In XX we plotted a Hill plot for differet k's where .....

```{r pressure, echo=FALSE, fig.cap="MSE for sample sizes n=30, 100, 200 for the tail and scale estimations via log-moments and maximum likelihood estimation", out.width = '100%'}
knitr::include_graphics("MSE_n25_100_200.pdf")
```
<!-- The QQ-estimator [@Kratz96] is closely related to  -->
<!-- the Hill estimator and fits a  -->
<!-- least squares line through the logarithms of the ordered statistics  -->
<!-- (y-axis) and the corresponding quantiles of the exponential distribution  -->
<!-- (x-axis). The reciprocal slope is returned as the estimate of the tail  -->
<!-- exponent. -->
<!-- For instance, the top-left panel in Figure \ref{fig:QQ-plots} shows the  -->
<!-- (static) QQ-estimator for the magnitudes of the solar flare data, returning  -->
<!-- a good fit to a Pareto distribution with tail parameter  -->
<!-- `r round(flares_exponent, 2)`. The dynamic QQ-estimator plot (top-center panel) -->
<!-- plots the tail exponent estimate for the largest values at cutoffs  -->
<!-- varying up to the 5th order statistic, and the region of stability at  -->
<!-- 0.8 indicates a recommended estimate. -->

<!-- However, as described nicely by @Resnick97, the less  -->
<!-- similar a heavy-tailed distribution is to a Pareto distribution, the less -->
<!-- useful a Hill or QQ-Plot estimator becomes.  -->
<!-- The bottom-left panel of Figure \ref{fig:QQ-plots}, for instance, shows the  -->
<!-- static QQ-estimator plot for $2000$ draws from the Mittag-Leffler distribution with  -->
<!-- tail parameter $0.8$. The stretched exponential shape means  -->
<!-- that the Mittag-Leffler distribution has more probability near $0$ than  -->
<!-- the Pareto distribution, severely biasing the estimator downwards  -->
<!-- (`r round(ML_exponent, 2)`).  -->
<!-- Even by looking at different cutoffs via the dynamic QQ-estimator plot  -->
<!-- (bottom-center panel) one hardly identifies $0.8$ as a clear candidate for  -->
<!-- a tail parameter estimate. -->

<!-- QQ-estimators, and the closely related Hill estimator, are hence not  -->
<!-- suitable to detect a heavy-tailed Mittag-Leffler distribution. Moreover,  -->
<!-- QQ-estimator plots of exponentially distributed data are virtually  -->
<!-- indistinguishable from the bottom-left panel. This shows that  -->
<!-- _Mittag-Leffler distributed data may look like exponentially distributed  -->
<!-- data if examined via a QQ-estimator_.  -->

```{r likelihood-ratio-test}
flares_arrivals <- flares %>% ctre() %>% thin(k = 500) %>% interarrival()
mlmle_out <- mlmle(data = flares_arrivals)
```
<!--Anpassen:
<!-- Hence if there is some prior expectation that the data are drawn from the  -->
<!-- Mittag-Leffler distribution -->
<!-- (as is the the case for threshold exceedance times), -->
<!-- then we recommend avoiding QQ-estimators and Hill plots altogether,  -->
<!-- and instead examining QQ-plots directly, on a logarithmic scale  -->
<!-- (see Figure \ref{fig:QQ-plots}, right column). -->
<!-- The scale parameter $\sigma$ is irrelevant for a QQ-Plot.  -->
<!-- The tail parameter $\beta$ can be estimated quickly via the log-moment  -->
<!-- method by @Cahoy2013, or via maximum likelihood.  -->
<!-- Both estimators are implemented in `MittagLeffleR`  -->
<!-- [@MittagLeffleR].  -->

Since the exponential distribution is nested in the Mittag-Leffler family of distributions, an appropriate way to choose between a model with exponential and Mittag-Leffler inter-exceedance times seems to be a Likelihood ratio test. Although the two models are nested, the assymptotic distribution seems not to be $\chi^2_1$ since under $H_0$ the paremter $\beta$ of Mittag-Leffler distribution is equal to zero which is on the boundary of the parameter space $(0,1]$.

@@

<!-- Anpassen:
<!-- As an example, the threshold crossing times for the solar flare dataset -->
<!-- (Figure \ref{fig:flare-diagnostics-2}, right panel) -->
<!-- yield .  -->

@@

* We choose $\ell$ from the order statistics, i.e. $\ell$ is the $k$-th
  largest of the observations $X_j$, where $k$ runs from 
  $k_\text{min}, k_\text{min} + 1, \ldots, k_\text{max}$. 
  The datasets are then of length $k-1$.
* We use $k$ rather than $\ell$ for the horizontal axis of our plots. 
* In Step 4, rather than plotting $p^{1/\hat \beta} \hat \sigma(\ell)$
  we plot $k^{1/\hat \beta} \hat \sigma(\ell)$. This changes 
  $\hat \sigma_0$ by the multiplicative constant $n^{1/\hat \beta}$, 
  but has the advantage that
  $\hat \sigma_0$ does not change if one pre-processes the data by 
  removing all observations below a certain threshold.

@@

# Checking Model Assumptions

```{r flare-diagnostics-2, fig.height=7, fig.cap="Diagnostic plots for the solar flare data: empirical copula and QQ Plot. \\label{fig:flare-diagnostics-2}"}
par(mfrow = c(2,2))
empcopula(flares_ctre)
flares_ctre %>% interarrival() %>% mlqqplot(log = 'xy', tail = 0.8, main = "Mittag-Leffler QQ Plot")
```

i.i.d.:
: After removing the "noise observations" below the smallest threshold 
  $\ell_0$, the pair sequence $(T(\ell_0, i), X(\ell_0,i))$ is i.i.d. 
  An indication if this is true is given 
  by an auto-correlation plot for the logarithms (to ensure finite moments) 
  of the two time series.

Uncoupled:
: Each $T(\ell, i)$ is independent of each $X(\ell, i)$. We propose an empirical copula
  plot to check for any dependence.

${\rm ML}(\beta, \sigma)$ distribution of $T(\ell, i)$:
: Apply a cutoff at the lowest threshold $\ell_0$, 
  extract the threshold crossing times, 
  and create a QQ Plot for the Mittag-Leffler distribution. 
  Use a log-Moment estimate of the tail parameter for the theoretical / population 
  quantiles of the plot.

@@

According to Figure \ref{fig:flares}, for a threshold $\ell$ at the $k$-th order 
statistic, the fitted threshold exceedance time distribution is 
$$
T_\ell \sim {\rm ML}(\beta, k^{1/\beta} \sigma_0), 
$$
where $\beta = 0.85$ and $\sigma_0 = 3.0 \times 10^7 {\rm sec}$. 
Unlike the exponential distribution, the Mittag-Leffler distribution is not 
memoryless, and the probability density of the time $t$ until the next threshold 
crossing will depend on the time $t_0$ elapsed since the last threshold crossing.
This density equals 
$$
p(t|\beta, \sigma_0, \ell, t_0) = \frac{f(t + t_0 | \beta, k^{1/\beta} \sigma_0)}{\mathbf P[T_\ell > t_0]}
$$
where $f(\,\cdot\, | \beta, k^{1/\beta} \sigma_0)$ is the probability density of 
${\rm ML}(\beta, k^{1/\beta} \sigma_0)$. 
The more time has passed without a threshold crossing, the more the probability 
distribution shifts towards larger values for the next crossing
(see Figure \ref{fig:hazard}, left panel). 
The hazard rate 
$$
h(t) = \frac{f(t| \beta, k^{1/\beta} \sigma_0))}{\int_t^\infty f(\tau| \beta, k^{1/\beta} \sigma_0))\,d\tau}
$$
represents the risk of a threshold crossing per unit time, and is a 
decreasing function for the Mittag-Leffler distribution. 
The closer $\beta$ is to $1$, the more the hazard rate mimics that of an 
exponential distribution (a constant function, see Figure \ref{fig:hazard}, 
right panel).

>>>>

---
title: "Statistical Inference for inter-arrival times of extreme events in bursty time series"
# titlerunning: "CTRE: POT for Bursty Time Series"
thanks: |
    Peter Straka was supported by the 
    Discovery Early Career Research Award DE160101147 on the Project 
    "Predicting Extremes when Events Occur in Bursts" by the Australian
    Research Council.
    Katharina Hees was supported by 
    the DAAD co-financed by the German Federal Ministry of Education and 
    Research (BMBF).

authors: 
- name: "Katharina Hees"
  address: TU Dortmund University
  email: hees@statistik.tu-dortmund.de

@@

bibliography: CTRMstats.bib
output: 
       rticles::elsevier_article:
                fig_caption: yes
                number_sections: true

@@

preamble: | 
  \usepackage{amssymb}
  \usepackage{natbib}

---

```{r get-rticles, eval=FALSE, include=FALSE, cache=TRUE}
# use this rticles version: 
devtools::install_github("rstudio/rticles")
# use the newest MittagLeffleR version:
devtools::install_github("UNSW-MATH/MittagLeffleR")
library(MittagLeffleR)
devtools::install_github("UNSW-MATH/CTRE")
library(CTRE)
setwd("~/GitHub/bursty-POT/article_springer")
Sys.setenv(TEXINPUTS=getwd(),
           BIBINPUTS=getwd(),
           BSTINPUTS=getwd())
```

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = FALSE,
  cache = TRUE,
  message = FALSE,
  fig.height = 3, out.width = '\\textwidth'
  )
library(CTRE)
library(MittagLeffleR)
library(magrittr)
set.seed(12345)
```

@@

We review the marked renewal process in Section 2, and 
derive a scaling limit theorem for inter-exceedance times in Section 3.
We give a statistical procedure to estimate model parameters via stability plots 
in Section 5, but to set the stage we first discuss inference for the 
Mittag-Leffler distribution in Section 4. In Section 6, we discuss the memory property of the 
Mittag-Leffler distribution, and how it affects the predictive distribution 
for the time until the next threshold crossing event. A simulation study of the effectiveness of our statistical procedure is given in Section 7. In Section 8 we apply our method to a real data set. Finally we close with a discussion and conclusion in Section 8.
For all statistical computations we have used R [@R]. 
All code and data used for the analysis in this article has been organized into 
an R package `CTRE` (<https://github.com/UNSW-MATH/CTRE>).
The source code for the figures generated in this manuscript is 
available online at <https://github.com/UNSW-MATH/bursty-POT>.

@@

In this section we will As a model for extreme observations, we use a Marked Renewal Process (MRP):

@@

: Let $(W,J), (W_1, J_1), (W_2, J_2), \ldots$ be i.i.d. pairs of random 
variables, where the $W_k > 0$ are interpreted as the *waiting times* 
and $J_k \in [x_L, x_R]$ as the *event magnitudes* 
($x_L \in [-\infty, +\infty), x_R \in (-\infty, +\infty]$). 
If $W$ and $J$ are independent, the Marked Renewal Process is said to be *uncoupled*.

@@

: Given a threshold $\ell \in (x_L, x_R)$, 
consider the stopping time 
$$\tau(\ell) := \min\{k: J_k > \ell\},\quad \ell \in (x_L, x_R).$$
Define the pair of random variables $(X(\ell), T(\ell))$ via 
$$X(\ell) = J_{\tau(\ell)} - \ell, \quad 
T(\ell) = \sum_{k=1}^{\tau(\ell)} W_k.$$
By restarting the MRP at $\tau(\ell)$, inductively define the
two i.i.d. sequences $T(\ell,n)$ and $X(\ell, n)$, $n \in \mathbb N$, 
called the "interarrival times" and the "exceedances", respectively.
The pair sequence $(T(\ell, n), W(\ell, n))_{n \in \mathbb N}$ is called 
a Continuous Time Random Exceedance model (CTRE). 
If the underlying MRP is uncoupled, then the CTRE is also called 
uncoupled.

```{r thresholdedBursty, message=FALSE, fig.height=7, fig.width=9, fig.cap="\\label{fig:thresholdedBursty}Exceedances (red) and Times until Exceedance (durations between blue crosses) for a given threshold $\\ell$ (dashed line)."}
tail <- params$tail
n <- params$n 
sigma <- (cos(pi*tail/2))^(1/tail)

times <- cumsum(stabledist::rstable(n, tail, 1, sigma, pm=1))
#times<-cumsum(rpareto(n,tail))
magnitudes <- extRemes::revd(n, scale = 1, shape = 0)
sim_ctre <- ctre(data.frame(times, magnitudes)) 
par(mfrow = c(2,1))
plot(sim_ctre, p = 0.01, main = "Simulated MRP") 
flares %>% ctre() %>% plot(p = 0.02, log = 'y', main = "HXRBS data") 
```

In this article, we restrict ourselves to the uncoupled case, where W and J are independent. Then the two sequences $X(\ell, n)_{n \in \mathbb N}$ and $T(\ell, n)_{n \in \mathbb N}$ are independent as well. To see why, note that $X(\ell)$ is, in distribution, simply equal to $J | J > \ell$, independent of any waiting time $W_k$.  
Figure \ref{fig:thresholdedBursty} shows a simulated dataset in the top panel, where $W$ has a stable distribution with tail parameter $\beta =$ `r params$tail` (and skewness $1$ and location $0$), and where $J$ is from a standard Gumbel distribution. In the bottom panel, we plot a time series of solar flare intensities derived from a NASA dataset [@HXRBS] which we will later examine more closely (see Section 8). Clearly, the simulated data exhibit long intervals 
_without any_ events, whereas in the real-world dataset events appear continuously. The threshold exceedances, however, appear to have similar statistical behaviour in both models. Observations below a threshold are commonly discarded in Extreme Value Theory (POT approach); likewise, the CTRE model interprets these observations as noise and filters them out.

@@

: Let the waiting times $J_k$ be in the domain of attraction of a positively skewed sum-stable law with stability parameter $0 < \beta < 1$; more precisely,
\begin{align} \label{eq:stability}
\frac{W_1 + \ldots + W_n}{b(n)} \overset{d}{\longrightarrow} D, 
\quad n \to \infty
\end{align}
for a function $b(n)$ which is regularly varying at $\infty$ with parameter $1/\beta$, and where $\mathbf E[\exp(-sD)] = \exp(-s^\beta)$. Write $p := \mathbf P(J > \ell)$. Then the weak convergence
$$
\frac{T(\ell)} {b(1/p)} \to W_\beta \quad \text{ as } \quad \ell \uparrow x_R
$$
holds, where the Mittag-Leffler random variable $W_\beta$ is defined on 
the positive real numbers via 
$$
\mathbf E[\exp(-sW_\beta)] = \frac{1}{1+s^\beta}.
$$

*Proof of Theorem:* We interpret the threshold crossing time $T(\ell)$ as the hitting time of the 
underlying CTRM (Continuous Time Random Maxima) or "max-renewal process", 
and then utilize a result by @MeerschaertStoev08. 
The running maximum process is defined as 
$$
M(c) := J_1 \vee \ldots \vee J_{\lfloor c \rfloor},
$$
and since we assume that the $J_k$ have a continuous distribution, there exist 
norming functions $a(c)$ and $d(c)$ such that 
$$
\mathbf P\left[ \frac{M(c) - d(c)}{a(c)} \le \ell^* \right] 
\longrightarrow F(\ell^*), \quad t \to \infty
$$
where $F$ is a generalized extreme value distribution, and $\ell^*$ is any 
value from the support of $F$. 
The CTRM process is then defined via 
$$
V(t) = M(N(t)), \quad t \ge 0
$$
where $N(t)$ is the renewal process associated with the waiting times 
$W_k$: 
$$
N(t) = \max\{n: W_1 + \ldots + W_n \le t\}.
$$
Now a key observation is that
$$
T(\ell) = \inf\{t: V(t) > \ell\}, 
$$
and that 
$$
T(\ell) > t \quad \text{ if and only if } \quad V(t) \le \ell.
$$
By [Theorem 3.1, @MeerschaertStoev08], we have the stochastic process 
convergence 
$$
\frac{V(ct) - d(\tilde b(c))}{a(\tilde b(c))} 
\stackrel{d}{\longrightarrow} Y(t), \quad t > 0.
$$
where $Y(t)$ is a time-changed ("subordinated") extremal process, and where 
$\tilde b(c)$ is a regularly varying norming function which is
_inverse_ to $b(c)$, in the sense that 
$b(\tilde b(c)) \sim c \sim \tilde b(b(c))$.

@@

For a scale parameter $\sigma > 0$, we write ${\rm ML}(\beta, \sigma)$ for the distribution of $\sigma W_\beta$. The Mittag-Leffler distribution with parameter $\beta \in (0,1]$ is 
a heavy-tailed positive distribution for $\beta < 1$, with infinite mean. 
However, as $\beta \uparrow 1$, ${\rm ML}(\beta, \sigma)$ converges 
weakly to the exponential distribution ${\rm Exp}(\sigma)$.
This means that although its moments are all infinite, the Mittag-Leffler
distribution may (if $\beta$ is close to 1) be indistinguishable from the 
exponential distribution, for the purposes of applied statistics. 
For a detailed reference on the Mittag-Leffler distribution, see e.g. 
@Haubold11, and for algorithms, see e.g. the R package `MittagLeffleR`
[@MittagLeffleR].

**Remark:**

: If $\beta = 1$, the result of the Theorem above is standard, see e.g. 
Equation (2.2) in @Gut1999.

@@

In the classical POT approach a GPD distribution is fitted to the data. Due to the Theorem in the last Section we know that for high threshholds the waiting times between the exceedances are Mittag-Leffler distributed in case of heavy tailed waiting times. Hence, we discuss in this section first parameter estimation for Mittag-Leffler distributions.

The first method proposed for parameter estimation of the Mittag-Leffler distribution was the fractional moment estimator by [@kozubowski2001]. One drawback of this method is that constant priors for the tail parameter are needed for the calculation of the estimates. [@Cahoy2010] proposed an estimator based on the log-Moments which doesn't need prior knowledge. Furthermore, they showed in a simulation study that the log-Moments estimator outclasses the fractional moment estimator regarding bias and root mean squared error (RMSE). Due to the form of the density function there exists no closed form for the Maximum Likelihood estimator (MLE). In the RPackage `MittagLeffleR` Maximum Likelihood estimation is implemented via numerical optimization. The MLE outperforms the log-moment estimator regarding bias and RMSE for big enough sample sizes, but is extremely computational intensive and the differences in bias and RMSE between the two methods are very small. In Figure \ref{fig:MSE} one can see that both methods work quite well even for very low sample sizes.    
Since the Mittag-Leffler distribution is heavy-tailed, many researchers would intuitively give the highest importance to the tail behaviour of the distribution. Of course, one can also use established tail exponent estimators such as the Hill estimator. However, these methods are less statistically efficient since they only use a portion of the information contained in the data, as was already mentionded by [@kozubowski2001]. Addtionally they need a specification of a tuning parameter. In XX one can see a Hill plot for differet k's.

```{r MSE, echo=FALSE, fig.height=12, fig.cap="\\label{fig:MSE} MSE for sample sizes n=30, 100, 200 for the tail and scale estimations via log-moments and maximum likelihood estimation", out.width = '100%'}
knitr::include_graphics("MSE_n25_100_200.pdf")
```

```{r likelihood-ratio-test}
flares_arrivals <- flares %>% ctre() %>% thin(k = 500) %>% interarrival()
mlmle_out <- mlmle(data = flares_arrivals)
```

Since the exponential distribution is nested in the Mittag-Leffler family of distributions, an appropriate way to choose between a model with exponential and Mittag-Leffler inter-exceedance times seems to be a Likelihood ratio test. Although the two models are nested, the assymptotic distribution is not $\chi^2_1$-distributed, the classical Wilks Theorem doesn't apply since under $H_0$ the paremter $\beta$ of Mittag-Leffler distribution is equal to zero which is on the boundary of the parameter space $(0,1]$. We used a bootstrapped likelihood ratio test. In figure XX one can see the power for different tail parameters. As expected the power gets smaller for tail parameters close to one, since the Mittag-Leffler distribution converges for $\beta \rightarrow 1$ to an exponetial distribution. For tail parameters close to one it is hard to differentiate a Mittag-Lefller distribution from a exponential distribution.

@@

@@

* We choose $\ell$ from the order statistics, i.e. $\ell$ is the $k$-th
  largest of the observations $X_j$, where $k$ runs from 
  $k_\text{min}, k_\text{min} + 1, \ldots, k_\text{max}$. 
  The datasets are then of length $k-1$.

* We use $k$ rather than $\ell$ for the horizontal axis of our plots.

* In Step 4, rather than plotting $p^{1/\hat \beta} \hat \sigma(\ell)$
  we plot $k^{1/\hat \beta} \hat \sigma(\ell)$. This changes 
  $\hat \sigma_0$ by the multiplicative constant $n^{1/\hat \beta}$, 
  but has the advantage that
  $\hat \sigma_0$ does not change if one pre-processes the data by 
  removing all observations below a certain threshold.

@@

@@

According to Figure \ref{fig:flares}, for a threshold $\ell$ at the $k$-th order 
statistic, the fitted threshold exceedance time distribution is 
$$
T_\ell \sim {\rm ML}(\beta, k^{1/\beta} \sigma_0), 
$$
where $\beta = 0.85$ and $\sigma_0 = 3.0 \times 10^7 {\rm sec}$. 
Unlike the exponential distribution, the Mittag-Leffler distribution is not 
memoryless, and the probability density of the time $t$ until the next threshold 
crossing will depend on the time $t_0$ elapsed since the last threshold crossing.
This density equals 
$$
p(t|\beta, \sigma_0, \ell, t_0) = \frac{f(t + t_0 | \beta, k^{1/\beta} \sigma_0)}{\mathbf P[T_\ell > t_0]}
$$
where $f(\,\cdot\, | \beta, k^{1/\beta} \sigma_0)$ is the probability density of 
${\rm ML}(\beta, k^{1/\beta} \sigma_0)$. 
The more time has passed without a threshold crossing, the more the probability 
distribution shifts towards larger values for the next crossing
(see Figure \ref{fig:hazard}, left panel). 
The hazard rate 
$$
h(t) = \frac{f(t| \beta, k^{1/\beta} \sigma_0))}{\int_t^\infty f(\tau| \beta, k^{1/\beta} \sigma_0))\,d\tau}
$$
represents the risk of a threshold crossing per unit time, and is a 
decreasing function for the Mittag-Leffler distribution. 
```{r hazard, fig.height=5, fig.cap="Left: Conditional distribution of time until next threshold crossing, depending on elapsed time $t_0$ since last crossing ($\\beta = 0.8$, $\\sigma_0 = 1$). Right: Hazard rate depending on tail parameter $\\beta$.\\label{fig:hazard}"}
tail <- 0.7
scale <- 1
t_0 <- c(0, 1, 10)
from <- 0.1
to <- 100 * scale
xx <- exp(seq(from = log(from), to = log(to), length.out = 100))
par(mfrow = c(1,2))
plot(xx, rep(1, length(xx)), ylim = c(0.001,1), type = 'n', log = 'xy', 
     xlab = 't', ylab = 'p')
for (k in 1:length(t_0)) {
  cond <- pml(q = t_0[k], tail = tail, scale = scale, lower.tail = FALSE)
  yy <- dml(x = xx + t_0[k], tail = tail, scale = scale) / cond
  lines(xx,yy, col = k+1, lty = k)
}
legend("bottomleft", c("t_0 = 0", "t_0 = 1", "t_0 = 10"), col = 2:4, lty = 1:3)
